---
title: "Visualization Practice"
author: "Christopher Junk"
date: "February 4, 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

First, I need to load in the sample data. 
```{r}
install.packages("gapminder")
library(gapminder)
library(ggplot2)
```
```{r}
gapminder
```

Assign the data to an object, p. 
```{r}
p <- ggplot(data = gapminder)
```

Need to tell ggplot which variable takes x and y. 
```{r}
p <-  ggplot(data = gapminder, 
             mapping = aes(x = gdpPercap, #This argument links variables to components of the plot. 
                           y = lifeExp))
p
str(p)
p + geom_point() # Up to now p was just a description of what goes where in the plot, the geom_ argument gives the plot its form. In this case it is a scatter plot. 
```

Now try the smoothed time series plot. 
```{r}
p + geom_smooth()
```
Now try the smoothed line with the dots as well. 
```{r}
p + geom_smooth() +
    geom_point()
```

Now try the linear model instead of the gam for smoothing. 
```{r}
p + geom_smooth(method = "lm") +
    geom_point()
```

Now change the X axis scale to natural log. 
```{r}
p + geom_smooth(method = "lm") +
    geom_point() + 
  scale_x_log10()
```

Look at the scales library and change the axis tick labels. 
```{r}
library(scales)
p + geom_smooth(method = "lm") +
    geom_point() + 
  scale_x_log10(labels = scales::dollar) #thelibrary::thefunction
```

Now let's try some options in the mapping argument. 
```{r}
p <- ggplot(data = gapminder, 
            mapping = aes(x = gdpPercap,
                         y = lifeExp, 
                         color = continent))

p + geom_point() +
    geom_smooth(method = "loess") + 
    scale_x_log10(labels = scales::dollar)
```

Now, get rid of that color and make all points purple. 
```{r}
p <- ggplot(data = gapminder, 
            mapping = aes(x = gdpPercap,
                         y = lifeExp,
                         color = continent))

p + geom_point(color = "purple") +
    geom_smooth(method = "loess") + 
    scale_x_log10(labels = scales::dollar)
```


Now adjust the line size and point size. 
```{r}
p <- ggplot(data = gapminder,
            mapping = aes(x = gdpPercap,
                          y = lifeExp)) 
p + geom_point(alpha = 0.3) +
    geom_smooth(color = "orange", se = FALSE, size = 8, method = "lm") +
    scale_x_log10()
```


Let's add some titles and axis labels. 
```{r}
p + geom_point(alpha = 0.3) +
    geom_smooth(method = "gam") +
    scale_x_log10(labels = scales::dollar) +
    labs(x = "GDP Per Capita", y = "Life Expectancy in Years",
         title = "Economic Growth and Life Expectancy",
         subtitle = "Data points are country-years",
         caption = "Source: Gapminder.")
```

Now they recreate what the first code would have been had they not intentionally made the purple mistake. 
```{r}
p <- ggplot(data = gapminder,
            mapping = aes(x = gdpPercap,
                          y = lifeExp,
                          color = continent))
p + geom_point() +
    geom_smooth(method = "loess") +
    scale_x_log10()
```

Fill in the confidence intervals. 
```{r}
p <- ggplot(data = gapminder,
            mapping = aes(x = gdpPercap,
                          y = lifeExp,
                          color = continent, 
                          fill = continent))
p + geom_point() +
    geom_smooth(method = "loess") +
    scale_x_log10()
```

Break down the code: change aesthetics in each geom. 
```{r}
p <- ggplot(data = gapminder, mapping = aes(x = gdpPercap, y = lifeExp))
p + geom_point(mapping = aes(color = continent)) +
    geom_smooth(method = "loess") +
    scale_x_log10()
```

Create a color gradient for a continuous variable: logged population total. 
```{r}
p <- ggplot(data = gapminder,
            mapping = aes(x = gdpPercap,
                          y = lifeExp))
p + geom_point(mapping = aes(color = log(pop))) +
    scale_x_log10()
```

Practice saving. 
```{r, fig.width=12, fig.height=9}
p_out <- p + geom_point() +
    geom_smooth(method = "loess") +
    scale_x_log10()

ggsave("my_figure.pdf", plot = p_out)
```

Play around with their suggestions. 
```{r}
p <- ggplot(data = gapminder,
            mapping = aes(x = gdpPercap,
                          y = lifeExp))
p + geom_smooth() + 
    geom_point() #The first geom is on the bottom, and it layers on top. 
```

Change the x to population. 
```{r}
p <- ggplot(data = gapminder,
            mapping = aes(x = pop,
                          y = lifeExp))
p + geom_smooth() + 
    geom_point()

p + geom_point() +
    geom_point() + 
    scale_x_sqrt() #this is nonsense, but I understand the way scale changes works now. 

p <- ggplot(data = gapminder,
            mapping = aes(x = pop,
                          y = lifeExp,
                          color = year))
p + geom_point() + 
    geom_smooth()

p <- ggplot(data = gapminder,
            mapping = aes(x = pop,
                          y = lifeExp,
                          color = factor(year)))
p + geom_point() + 
    geom_smooth()
```

